---
title: "williams-soilhydro-option-a"
output: html_document
date: "2024-11-18"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
list.files()
# install.packages('soilDB', dependencies = TRUE)
# install.packages("readr")
library(tidyverse)
library(soilDB)
```

```{r}
testdata <- read_csv("data/TESTSoil_data - Sheet1.csv")
```

Creating a tibble for the first layer `neiu	1	0-10`

```{r trying shamyas model}
test <- tibble(
  sand = c(0),
  silt = c(86.4),
  clay = c(13.6)
)
test
```

Create variable names vector for `neiu	1	0-10`

```{r trying shamyas model}
vars <- c("sand", "silt", "clay")
vars
```

```{r try-manual-rosetta-function}
test1 <- ROSETTA(test, vars, v = "3")
test1
```

## Trying Gavin's Option A

```{r bones-of-option-a}
# option A
testdata %>% 
  mutate(new_variable = function(existing_var, ...))
```

```{r}
optiona <- testdata %>%
  mutate(rosetta_vars = ROSETTA(., vars = c("sand_percent", "silt_percent", "clay_percent"), v = "3"))
```
I think it worked!

## FROM GAVIN: Great job! Within the `ROSETTA()` function, I don't think you need `x`, because you're already working within testdata, for "vars" you want to specify 'sand', 'silt', 'clay'. I tried that but it was still giving more output than is necessary. I thin we can simply input testdata as 'x'.

```{r simpler-version}
ROSETTA(testdata, vars = c("sand_percent", "silt_percent", "clay_percent"), v = "3")
```

## Gavin again:: Yay I think it worked!

## important to note that "test1" is actually just neiu 0-10... im not sure if everything I want is included tho (gives rosetta_vars as data.frame 1x 17)... also not sure if the info in the new column is just a repeat of row 1 ???

```{r inspecting?}
optiona$"rosetta_vars"
```
## has two sets of theta r, theta s, alpha, and n par...

```{r}
optiona
```

```{r}
soildata <- read_csv("Soil_data - Data Sheet.csv")
soildata
```

```{r}
soildata2 <- ROSETTA(soildata, vars = c("obs_sand_pct", "obs_silt_pct", "obs_clay_pct"), v = "3")
soildata2
```


```{r neiu data graphs, fig.height=5, fig.width=5}
neiu_graph <- soildata2 %>% 
  filter(core_name == "neiu") %>% 
  mutate(depth_increment = as.numeric(substr(core_depth, 1, 1))) %>% 
    ggplot(aes(x = -depth_increment*10-5, y = ksat, color = core_no., shape = core_no.)) +
  geom_line() +
  geom_point(size = 2) +   
  scale_color_manual(values = c(`1` = "maroon", `2` = "darkmagenta", `3` = "deeppink1")) +
  theme_minimal() +
  scale_x_continuous(limits = c(-100, 0)) +
  scale_y_continuous(limits = c(0.75, 3.25)) + 
  labs(title = "Hyrdaulic Conductivity Comparisons by Depth, NEIU", x = "Depth, cm", y = "Hydraulic Conductivity, cm/day") +
  coord_flip()
neiu_graph
```

```{r csu graph, fig.height=5, fig.width=5}
csu_graph <- soildata2 %>% 
  filter(core_name == "csu") %>% 
  mutate(depth_increment = as.numeric(substr(core_depth, 1, 1))) %>% 
    ggplot(aes(x = -depth_increment*10-5, y = ksat, color = core_no., shape = core_no.)) +
  geom_line() +
  geom_point(size = 2) + 
  scale_color_manual(values = c(`1` = "coral", `trees` = "brown"), 
                     labels = c("1", "2")) +
  theme_minimal() +
  scale_x_continuous(limits = c(-100, 0)) +
  scale_y_continuous(limits = c(0.75, 3.25)) +
  labs(title = "Hyrdaulic Conductivity Comparisons by Depth, CSU", x = "Depth, cm", y = "Hydraulic Conductivity, cm/day") +
  coord_flip()
csu_graph
```

```{r uic data graphs, fig.height=6, fig.width=6}
uic_graph <- soildata2 %>% 
  filter(core_name == "uic") %>% 
  mutate(depth_increment = as.numeric(substr(core_depth, 1, 1))) %>% 
    ggplot(aes(x = -depth_increment*10-5, y = ksat, color = core_no., shape = core_no.)) +
  geom_line() +
  geom_point(size = 2) + 
  scale_color_manual(values = c(`GH Ne` = "purple", `GH Pr` = "mediumorchid1", `GH Sw` = "plum2", `GH Veg` = "palevioletred")) +
  theme_minimal() +
  scale_x_continuous(limits = c(-100, 0)) +
  labs(title = "Hyrdaulic Conductivity Comparisons by Depth, UIC Greenhouse", x = "Depth, cm", y = "Hydraulic Conductivity, cm/day") +
  coord_flip()
uic_graph
```

```{###r combined graphs NOT USING}
combined_data <- soildata2 %>%
  filter(core_name %in% c("neiu", "csu", "uic")) %>%
  mutate(depth_increment = as.numeric(substr(core_depth, 1, 1)))

final_plot <- ggplot(combined_data, aes(x = -depth_increment*10-5, y = ksat, color = core_name)) +
  geom_line() + 
  scale_color_manual(values = c("neiu" = "red", "csu" = "blue", "uic" = "green")) +  
  scale_x_continuous(limits = c(-100, 0)) +
  labs(title = "Hydraulic Conductivity Comparisons by Depth",
       x = "Depth, cm",
       y = "Hydraulic Conductivity, cm/day",
       color = "Core Sites") +
  coord_flip()

final_plot
```

```{r tidying for theta graphs}
neiu_soildata <- soildata2 %>% filter(core_name == "neiu")
neiu_soildata

csu_soildata <- soildata2 %>% filter(core_name == "csu")
csu_soildata

uic_soildata <- soildata2 %>% filter(core_name == "uic")
uic_soildata
```


```{r soil capacity and residual graphs}
neiu_soildata_long <- neiu_soildata %>% 
  pivot_longer(cols = c(theta_r, theta_s), 
                          names_to = "theta", 
                          values_to = "value") 
neiu_soildata_long


neiu_soil_cr <- neiu_soildata_long %>% 
  filter(core_name == "neiu") %>% 
  mutate(theta = fct_rev(theta), 
         core_depth = fct_rev(core_depth)) %>% 
    ggplot(aes(x = core_depth, y = value*100, fill = theta)) +
  geom_bar(stat = "identity", width = 0.5) +  
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  scale_fill_manual(values = c("theta_r" = "deepskyblue4", "theta_s" = "deepskyblue1"), 
                    labels = c("Soil Capacity", "Soil Residual")) +
  labs(title = "Volumetric Soil Water Capacity and Residual by Depth, NEIU",
       x = "Depth, cm",
       y = " Volumetric Water Capacity and Residual, %",
       fill = "Legend") +  
  facet_wrap(~core_no., ncol = 3) +
  theme_minimal() +
  coord_flip()

neiu_soil_cr
```

```{r soil capacity and residual graphs csu}
csu_soildata_long <- csu_soildata %>% 
  pivot_longer(cols = c(theta_r, theta_s), 
                          names_to = "theta", 
                          values_to = "value") 
csu_soildata_long


csu_soil_cr <- csu_soildata_long %>% 
  filter(core_name == "csu") %>% 
  mutate(theta = fct_rev(theta), 
         core_depth = fct_rev(core_depth)) %>% 
    ggplot(aes(x = core_depth, y = value*100, fill = theta)) +
  geom_bar(stat = "identity", width = 0.5) +  
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  scale_fill_manual(values = c("theta_r" = "deepskyblue4", "theta_s" = "deepskyblue1"), 
                    labels = c("Soil Capacity", "Soil Residual")) +
  labs(title = "Volumetric Soil Water Capacity and Residual by Depth, CSU",
       x = "Depth, cm",
       y = " Volumetric Water Capacity and Residual, %",
       fill = "Legend") +  
  facet_wrap(~core_no., ncol = 3) +
  theme_minimal() +
  coord_flip()

csu_soil_cr
```

```{r soil capacity and residual graphs uic}
uic_soildata_long <- uic_soildata %>% 
  pivot_longer(cols = c(theta_r, theta_s), 
                          names_to = "theta", 
                          values_to = "value") 
uic_soildata_long


uic_soil_cr <- uic_soildata_long %>% 
  filter(core_name == "uic") %>% 
  mutate(theta = fct_rev(theta), 
         core_depth = fct_rev(core_depth)) %>% 
    ggplot(aes(x = core_depth, y = value*100, fill = theta)) +
  geom_bar(stat = "identity", width = 0.5) +  
  scale_y_continuous(breaks = seq(0, 100, by = 10)) +
  scale_fill_manual(values = c("theta_r" = "deepskyblue4", "theta_s" = "deepskyblue1"), 
                    labels = c("Soil Capacity", "Soil Residual")) +
  labs(title = "Volumetric Soil Water Capacity and Residual by Depth, UIC",
       x = "Depth, cm",
       y = " Volumetric Water Capacity and Residual, %",
       fill = "Legend") +  
  facet_wrap(~core_no., ncol = 4) +
  theme_minimal() +
  coord_flip()

uic_soil_cr
```





